Symmetric Circuits for Rank Gates

### 2018-04-08 sections/conclusion.tex
re-worked conclusions

\documentclass[../paper.tex]{subfiles}
\begin{document}

$\FPR$ is one of the two candidate logics for capturing $\PT$ an so
understanding its expressive power is an absolutely central problem in the
field. The main result of this paper establishes an equivalence between the
expressive power of $\FPR$ and the computational power of $\PT$-uniform families
of transparent symmetric rank-circuits. Not only does this result establish an
interesting characterisation of a little-understood and crucially important
logic, but it also further develops our understanding of the connection between
logic and circuit complexity, allowing our understanding of $\FPR$ (and
circuits) to benefit from ideas in both of these quite disparate fields.

This alternative view of the logic also helps emphasise certain characteristics
of the logic that may be much less obvious when working with the conventional
definition. To choose one notable example, the circuit characterisations (both
of $\FPR$ and $\FPC$) help bring to the fore the importance of the symmetry
properties inherent in the the syntactic structure of the logic. Indeed, we
recall that $\PT$-uniform families of transparent invariant (rather than
symmetric) rank-circuits and $\PT$-uniform families of invariant circuits with
symmetric gates decide exactly those queries that are in $\PT$. In this way the
restriction from families of invariant circuits to families of symmetric
circuits corresponds to the restriction from $\PT$ to $\FPR$ or $\FPC$. In this
sense, we might say that it is this underlying symmetry property of the logic,
brought into focus by the circuit characterisation, that separates $\FPC$ from
$\PT$, and which determines the relationship between $\FPR$ and $\PT$. In fact,
we have that $\FPR$ captures $\PT$ if, and only if, $\PT$-uniform families of
symmetric rank-circuits have the same computational power as $\PT$-uniform
families of invariant rank-circuits. In this way the circuit-characterisation
brings out this underlying symmetry property of the logic, demonstrates how
crucial this symmetry property for understanding the logic, and then provides us
with a framework for studying symmetry.

@@

However, much of the contribution of this paper is not in the main result
itself, but rather in the entirely novel framework for circuits and the numerous
new results that we needed to develop in order to prove the main result of this
paper. As such, we shall briefly review some of these contributions, placing an
emphasis both on the nature of these contributions and on the necessity of their
introduction. Indeed, it would seem at first glance that generalising Anderson
and Dawar's circuit characterisation of $\FPC$ to $\FPR$ should be a simple
matter of extending the model so as to allow circuits to be defined over bases
that includes rank-threshold functions -- and perhaps extending a few key
results and definitions as needed. We showed in
Theorem~\ref{thm:symmetric-circuits-bound} that, in fact, we cannot increase the
expressive power of the model by adding symmetric functions to the basis, and so
the addition of non-symmetric functions to the basis seems almost unavoidable.
However, while the inclusion of rank-threshold functions in the basis may seem
benign, it in fact violates a near-ubiquitous assumption in circuit complexity,
namely that the basis over which the circuit is defined must consist of
symmetric functions. While this assumption is usually not stated, the fact that
the circuit is taken to be a directed acyclic graph means that there is no order
(or any structure at all) on the inputs of each gate, an assumption which only
makes sense if it is assumed that the evaluation of each gate is invariant under
any ordering of its inputs -- which is to say the basis must be symmetric.
However, we found that in Anderson and Dawar's case this assumption pervades not
just their definition of a circuit, but also the central ideas of their paper,
including their definition of a circuit automorphism and of a symmetric circuit,
as well as almost all of the key results and techniques they use to establish
their characterisation.

As such, in order to include rank-threshold functions in the basis we needed to
construct a new framework for circuits as well as develop novel approaches and
techniques in order to work with these circuits. We began by generalising the
notion of a symmetric function, introducing isomorphism-invariant structured
functions. These functions take in structures encoded and strings and are
invariant under isomorphism, rather than under arbitrary permutations of the
string. It is easy to see that rank-threshold functions are
isomorphism-invariant. We defined a basis to be a set of isomorphism-invariant
functions, and extended the circuit model, adding to each gate in the circuit a
structure on its input gates in accord with the isomorphism-invariant structured
function labelling that gate. Rather than a gate being invariant under arbitrary
permutations of its input gates, a gate in this circuit should be invariant
under permutations of its input gate that form a (labelled) isomorphism. We
extended the notion of an automorphism to reflect this and defined what it means
for a circuit to be symmetric.

While we have generalised a few results for this framework, including the
support theorem, it is important to note that a number of crucial counting
techniques and polynomial-time algorithms for circuits are either not easy to
generalise or otherwise resist generalisation entirely. This follows from the
fact that a circuit automorphism needs to preserve more structure when acting on
non-symmetric gates than when acting on symmetric gates. As a result, while many
symmetry-related circuit properties are polynomial-time decidable for circuits
with symmetric gates, we have shown that these same properties are at least as
hard as the graph-isomorphism for circuits with non-symmetric gates. This is a
problem as the polynomial-time decidability of these properties is essential for
our translation from circuits into formulas.

% used throughout this paper, and is needed in order to apply the support
% theorem or reason about orbits, we need to

% for are we have shown that many natural circuit properties which are known to
% be polynomial-time decidable for circuits over symmetric bases are at least as
% hard to decide for circuits over non-symmetric bases as the graph isomorphism
% problem.

% This follows from in a rather subtle way from the fact that
% syntactic-equivalence for symmetric gates requires nothing more than checking
% if two quotiented sets are equal, while syntactic-equivalence between
% non-symmetric gates involves checking if two quotiented structures are equal.
% As such, while determining many

% While for circuits with functions into the basis. In Anderson and Dawar's case
% establishing the image of a gate under the action of an automorphism is
% entirely determined by the

% In particular, in order to define the translation from families of circuits to
% formulas we need that a number However, in order to prove this result we need
% to be able to compute in polynomial-time many basic symmetry-related
% properties of the circuit, and most of these properties reference the
% existence of an isomorphism between the structures associated with each gate.
% We show in Section~\ref{sec:transparent} that, in fact, deciding many of these
% properties for (general) rank-circuits is at least as hard as the
% graph-isomorphism problem. Importantly, the circuits studied by Anderson and
% Dawar have all of the relevant algorithmic, as in this case the isomorphism
% problems that arises reduces to set-equivalence, which is obviously
% polynomial-time decidable.

% In partoTo be more procides This is a problem, as these algorithmic properties
% are needed throughout the paper

% In particular, while almost all of the relevant properties are polynomial-time
% decidable for circuits with symmetric gates, enabling a in polynomial-time
% transformed into a normal form with unique extensions. Moreover, this normal
% form also many other crucial algorithmic properties needed to prove the main
% result, particularly the ability to compute the action of a permutation on a
% circuit, in polynomial-time. We showed in Section~\ref{} that these useful
% algorithmic properties are not (necessarily) preserved by the generalisation.
% In particular, we showed that none of these properties are polynomial-time
% decidable for rank-circuits (or any class of circuits over a non-symmetric
% basis) unless the graph isomorphism is in $\PT$.

% In order to prove our main result we $\PT$-uniform families of symmetric
% rank-circuits into equivalent formulas of $\FPR$ and visa

% Dawar. However, However, It would be natural to take symmetric rank-circuits
% as the natural generalisation of Anderson and Dawar's symmetric circuits and
% aim to prove an equivalence between the expressive power $\FPR$ and the
% computational power of $\PT$-uniform families symmetric rank-circuits.
% However, in order to prove this result we need to be able to compute in
% polynomial-time many basic symmetry-related properties of the circuit, and
% most of these properties reference the existence of an isomorphism between the
% structures associated with each gate. We show in Section~\ref{sec:transparent}
% that, in fact, deciding many of these properties for (general) rank-circuits
% is at least as hard as the graph-isomorphism problem. Importantly, the
% circuits studied by Anderson and Dawar have all of the relevant algorithmic,
% as in this case the isomorphism problems that arises reduces to
% set-equivalence, which is obviously polynomial-time decidable.

% In order to ensure

% when we introduce rank-gates this isomorphism problem that computing the
% action of an automorphism, or showing that a circuit is symmetric, or
% computing a great number of important properties is at least as hard for
% rank-circuits as the graph-isomorphism problem.

% to be able to apply the support theorem to circuits, and so needed a normal
% form for circuits with unique extensions, and we also needed to ensure that a
% number of basic circuit properties, and particularly those relating to
% symmetry, are polynomial-time decidable to circuits as well as compute a
% number of basic circuit properties in polynomial-time. This is not a problem
% for Anderson and Dawar, as all of the relevant circuit properties are
% polynomial-time decidable for symmetric-circuits with symmetric gates.
% Moreover, can be easily transformed into a normal form that has unique
% extensions (allowing for the application of the support theorem) and all of
% the nessasary circuit-properties are

% apply the support theorem, a crucial step in out proof, and in order to

% the introduction of structured functions complicates the analysis of such
% circuits. As we showed in Section~\ref{sec:transparent}, most of the
% properties of these rank-circuits are at least as hard to decide as the graph
% isomorphism problem. This poses a problem as the approach we take (and the
% approach Anderson and Dawar take) to proving this characterisation make
% repeated use of the polynomial-time decidability of these properties. Indeed,
% the two key results for translating $\PT$-uniform circuit families into
% equivalent formulas, the support theorem and the inductive evaluation of a
% gate (see Section~\ref{sec:circuits-to-formulas}), make use of polynomial-time
% algorithms for transforming a circuit into an equivalent circuit with unique
% labels, as well as for deciding if a circuit is symmetric, computing orbits,
% and deciding syntactic-equivalence.

As such, instead of working with general rank-circuits we have restricted our
attention to \emph{transparent} circuits, and proved our characterisation in
terms of $\PT$-uniform families of transparent symmetric rank-circuits. As we
have shown in Section~\ref{sec:transparent}, all of the relevant properties are
polynomial-time decidable for transparent circuits. Since all circuits with
symmetric gates are transparent, we might think of transparent circuits as an
alternative, more `algorithmically accessible', generalisation of the circuits
studied by Anderson and Dawar. However, while this restriction to transparent
circuits aids in our translation from circuits to formulas, it complicates our
translation in the other direction. Indeed, it is not hard to see that the
conventional approach for translating formulas of $\FO$ (or $\FPC$, or other
extensions of $\FO$) into $\PT$-uniform families of circuits does not, in
general, produce a family of transparent circuits. In order to address this we
defined a novel translation from formulas of $\FPR$ into equivalent
$\PT$-uniform families of transparent symmetric rank-circuits. This translation
is defined in two steps. First, we defined a translation from $\FPR$ to
$\PT$-uniform families of bounded-width formulas of $\FOrk$ and, second, we
defined a translation from these families of formulas into $\PT$-uniform
families of transparent symmetric rank-circuits. This translation, along with
the main result of the paper, gives us an equivalence between the expressive
power of three formalisms, which we summarise as follows:

@@

\begin{center}
  \begin{tikzcd}
    \FPR \ar[r, equal] \ar[dr, equal] & \parbox{0.35\textwidth}{$\PT$-uniform
      families of
      bounded-width $\FOrk$-formulas}  \ar[d, equal]\\
    & \parbox{0.38\textwidth}{$\PT$-uniform families of transparent symmetric
      rank-circuits}
  \end{tikzcd}
\end{center}

We already have a very similar result from from Anderson and
Dawar~\cite{AndersonD17} and Otto~\cite{Otto1996}:

\begin{center}
  \begin{tikzcd}
    \FPC \ar[r, equal] \ar[dr, equal] & \parbox{0.35\textwidth}{$\PT$-uniform
      families of
      bounded-width $\FOc$-formulas}  \ar[d, equal]\\
    & \parbox{0.35\textwidth}{$\PT$-uniform families of symmetric circuits with
      symmetric gates}
  \end{tikzcd}
\end{center}

These two sets of equivalences each give us three distinct formalisms equivalent
to $\FPR$ and $\FPC$, respectively. Importantly, each of these three formalisms
for $\FPC$ may be extended by a rank in quite distinct ways. We can extend
$\FOc$ by rank quantifiers and define $\FOrk$, we can add rank gates to
symmetric circuits with symmetric gates and define transparent symmetric
rank-circuits, and we can extend $\FPC$ by a rank operator and define $\FPR$. We
might expect these three quite different extensions to result in three
formalisms with quite different expressive power. However, the main result of
this paper along with the translation we discussed above, gives us that these
three extensions are indeed all equivalent. This provides evidence in favour of
the proposition that $\FPR$ is indeed the canonical, and most natural, extension
of $\FPC$. This is in contrast with the other rank logic in the literature, the
one introduced by Dawar et al.~\cite{Dawar09logicswith}, which was shown by
Gr\"{a}del and Pakusa~\cite{GradelP15a} to be strictly less expressive than
$\FPR$, and hence less expressive than $\PT$-uniform families of bounded-width
$\FOrk$ and $\PT$-uniform families of transparent symmetric rank-circuits.

@@

\subsection{Future Work}
We have shown that $\FPR$ has the same expressive power as $\PT$-uniform
families of transparent symmetric rank-circuits. It is easy to see that
$\PT$-uniform families of symmetric rank-circuits must be at least as powerful
as $\PT$-uniform families of transparent symmetric rank-circuits. It is natural
to ask if these two formalisms actually have the same expressive power or
whether the transparency restriction does indeed restrict the expressive power
of the model. It is worth noting that if families of symmetric rank-circuits and
families of transparent symmetric-rank circuits do not have the same expressive
power then $\FPR$ does not capture $\PT$.

In this paper we have developed a framework for working with circuits over
non-symmetric bases. In fact, the first time we explicitly use make any
reference to linear algebra is in the proof of
Lemma~\ref{lem:rank-triple-equivilence}, right towards the end of the paper. It
would be natural to ask if our circuit characterisation of $\FPR$ can be
generalised in order to develop similar characterisations of other logics. It
might be good idea to start with some other extension of $\FP$ by some operator
and try to develop a similar circuit characterisation in terms $\PT$-uniform
transparent symmetric circuits over an appropriate basis. There are numerous
logics that are similar in definition to $\FPR$ but which are not known to have
the same expressive power as $\FPR$, for example $\FP$ extended by an operator
that expresses the solubility of systems of equations over rings (rather than
fields). A first step might be to develop a circuit characterisation of one of
these logics.

@@

While we have developed a new and much more general circuit framework we have
not developed any approaches for proving lower bounds. As a first step out might
try to prove lower bounds for symmetric circuits with gates with bounded fan-in.
We might ask if it is possible to compute $\AND[3]$ using a symmetric circuit
with gates that have fan-in two.

An alternative, and perhaps more sophisticated, approach would be to try and use
pebble games, a standard tool for proving inexpressibility results in logic, to
directly prove lower bounds on circuits. Dawar~\cite{Dawar2016} has shown, via a
circuit-based argument, that the pebble games of Hella~\cite{Hella19961} can
played directly on symmetric circuits with symmetric gates without any need to
reference the logic. Dawar and Holm~\cite{DawarH2012} have defined a similar
pebble game for $\FPR$, and it would be interesting to know if an analogous
result can be established, allowing us play this pebble game directly on
symmetric rank-circuits and use it to prove lower bounds for these circuits.

We have shown that the $\PT$-uniform families of transparent symmetric
rank-circuits (resp. circuits with symmetric gates) have the same expressive
power as $\FPR$ (resp. $\FPC$), but we recall that if we remove the symmetry
restriction these circuit models express exactly those queries in $\PT$. A
natural next step, also suggested by Anderson and Dawar, would be try and weaken
the symmetry condition. One way this might be done would be to require that only
permutations from a particular subgroup of the symmetric group extend to
automorphisms of the circuit. It is easy to see that if we take this subgroup to
be the trivial subgroup then this nullifies the the symmetry requirement, and
the circuit model has the same expressive power as $\PT$, while if we take this
subgroup to be the entire symmetric group then we get back the full symmetry
requirement and hence our model has the same expressive power as the logic. As
such, this approach can be used to interpolate between the logic and $\PT$.

There is a rather easy way to develop a circuit characterisation for any logic
or complexity class with complete problems under first-order reductions. All
that is required is to extend the basis with the family of functions that decide
a complete problem and consider $\PT$-uniform families of transparent symmetric
circuits over this basis. We did not take this approach in our characterisation
of $\FPR$, preferring to use a more `natural' basis that more closely resembles
the actual definition of the logic. We should like to ask if there is a similar
circuit characterisation of $\CPTC$. One might start with bounded-rank fragments
of $\CPTC$ and see if these programs can be translated into circuits. We should
note that it does seem as though the symmetry inherent in $\CPTC$ programs is in
some sense different from the symmetry of symmetric circuits. It might be more
natural then to ask if $\CPTC$ can be characterised using families of circuits
with a weakened symmetry requirement.

>>>>

\documentclass[../paper.tex]{subfiles}
\begin{document}
$\FPR$ is one of the most expressive logics we know that is still
contained in $\PT$ and understanding its expressive power is an
important question.  The main result of this paper establishes an equivalence between the
expressive power of $\FPR$ and the computational power of uniform families
of transparent symmetric rank-circuits.  Not only does this result establish an
interesting characterisation of an important
logic, but it also deepens our understanding of the connection between
logic and circuit complexity in general and casts new light on
foundational aspects of the circuit model.

The circuit characterisation also helps emphasise certain important
aspects of the logic.  Given that $\PT$-uniform families of invariant
circuits (without the restriction to symmetry) express all properties
on $\PT$, we can understand  the inability of $\FPC$ (and,
conjecturally, $\FPR$) to express all such properties as essentially
down to symmetry.  As with other (machine) models of computation, the
translation to circuits exposes the inherent combinatorial structure
of an algorithm.  In the case of logics, we find that a key property
of this structure is its symmetry and the translation to circuits
provides us with the tools to study it.

% This alternative view of the logic also helps emphasise certain characteristics
% of the logic that may be much less obvious when working with the conventional
% definition. To choose one notable example, the circuit characterisations (both
% of $\FPR$ and $\FPC$) help bring to the fore the importance of the symmetry
% properties inherent in the the syntactic structure of the logic. Indeed, we
% recall that $\PT$-uniform families of transparent invariant (rather than
% symmetric) rank-circuits and $\PT$-uniform families of invariant circuits with
% symmetric gates decide exactly those queries that are in $\PT$. In this way the
% restriction from families of invariant circuits to families of symmetric
% circuits corresponds to the restriction from $\PT$ to $\FPR$ or $\FPC$. In this
% sense, we might say that it is this underlying symmetry property of the logic,
% brought into focus by the circuit characterisation, that separates $\FPC$ from
% $\PT$, and which determines the relationship between $\FPR$ and $\PT$. In fact,
% we have that $\FPR$ captures $\PT$ if, and only if, $\PT$-uniform families of
% symmetric rank-circuits have the same computational power as $\PT$-uniform
% families of invariant rank-circuits. In this way the circuit-characterisation
% brings out this underlying symmetry property of the logic, demonstrates how
% crucial this symmetry property for understanding the logic, and then provides us
% with a framework for studying symmetry.

@@

Still, the most significant contribution of this paper is not in the main result
but in the techniques that are developed to establish it, and we
highlight some of these now.  The
conclusion of~\cite{AndersonD17} says that the support theorem is
``largely agnostic to the particular [\ldots] basis'', suggesting that
it could be easily adapted to include other gates.  This turns out to
have been a misjudgment.  Attempting to prove the support theorem for
a basis that includes rank threshold gates showed us the extent to
which both the proof of the theorem and, more broadly, the definitions
of circuit classes, rest heavily on the assumption that all functions
computed by gates are symmetric.  Thus, in order to define what the
``symmetry'' condition might mean for circuits that include rank
threshold gates, we radically generalise the circuit framework to allow
for gates that take structured inputs (rather than sets of $0$s
and $1$s) and are invariant under isomorphisms.  This leads to a
refined notion of circuit automorphism, which allows us to formulate a
notion of symmetry and prove a version of the support theorem.  Again,
in that proof, substantial new methods are required.

% itself, but rather in the entirely novel framework for circuits and the numerous
% new results that we needed to develop in order to prove the main result of this
% paper. As such, we shall briefly review some of these contributions, placing an
% emphasis both on the nature of these contributions and on the necessity of their
% introduction. Indeed, it would seem at first glance that generalising Anderson
% and Dawar's circuit characterisation of $\FPC$ to $\FPR$ should be a simple
% matter of extending the model so as to allow circuits to be defined over bases
% that includes rank-threshold functions -- and perhaps extending a few key
% results and definitions as needed. We showed in
% Theorem~\ref{thm:symmetric-circuits-bound} that, in fact, we cannot increase the
% expressive power of the model by adding symmetric functions to the basis, and so
% the addition of non-symmetric functions to the basis seems almost unavoidable.
% However, while the inclusion of rank-threshold functions in the basis may seem
% benign, it in fact violates a near-ubiquitous assumption in circuit complexity,
% namely that the basis over which the circuit is defined must consist of
% symmetric functions. While this assumption is usually not stated, the fact that
% the circuit is taken to be a directed acyclic graph means that there is no order
% (or any structure at all) on the inputs of each gate, an assumption which only
% makes sense if it is assumed that the evaluation of each gate is invariant under
% any ordering of its inputs -- which is to say the basis must be symmetric.
% However, we found that in Anderson and Dawar's case this assumption pervades not
% just their definition of a circuit, but also the central ideas of their paper,
% including their definition of a circuit automorphism and of a symmetric circuit,
% as well as almost all of the key results and techniques they use to establish
% their characterisation.

% As such, in order to include rank-threshold functions in the basis we needed to
% construct a new framework for circuits as well as develop novel approaches and
% techniques in order to work with these circuits. We began by generalising the
% notion of a symmetric function, introducing isomorphism-invariant structured
% functions. These functions take in structures encoded and strings and are
% invariant under isomorphism, rather than under arbitrary permutations of the
% string. It is easy to see that rank-threshold functions are
% isomorphism-invariant. We defined a basis to be a set of isomorphism-invariant
% functions, and extended the circuit model, adding to each gate in the circuit a
% structure on its input gates in accord with the isomorphism-invariant structured
% function labelling that gate. Rather than a gate being invariant under arbitrary
% permutations of its input gates, a gate in this circuit should be invariant
% under permutations of its input gate that form a (labelled) isomorphism. We
% extended the notion of an automorphism to reflect this and defined what it means
% for a circuit to be symmetric.

While the new framework allows us to reproduce some results such as
the support theorem, something is lost in the generalisation.  When
our basis contains only symmetric functions, important properties of
a circuit, including the symmetry of the circuit itself, are easily
decidable.  In the new framework, checking circuit automorphism
requires checking isomorphism of structured inputs to gates and codes
hard problems.  For this reason, we imposed a further restriction to
the circuits in the form of \emph{transparency}.  But, it should be
clear that this is a restriction only in the broader framework.  All
symmetric circuits with only symmetric gates are transparent.  Thus,
transparent, symmetric circuits are still a generalisation of those
considered in~\cite{AndersonD17}.

The condition of transparency makes the translation of uniform circuit
families into formulas of logic (which is the difficult direction of
our characterisation) possible, but it complicates the other direction.
Indeed, the natural translation of formulas of $\FPR$ into uniform
circuit families yields circuits which are symmetric, but not
transparent.  This problem is addressed by introducing gadgets in the
translation---which for ease of exposition, we did in formulas of
$\FOrk$ which are then translated into circuits in the natural way.
Thus, the restriction to transparent circuits is sufficient to get
both directions of the characterisation.  Moreover, we argue that
something like it is necessary in order to avoid having to solve hard
isomorphism problems in the translation.

% While we have generalised a few results for this framework, including the
% support theorem, it is important to note that a number of crucial counting
% techniques and polynomial-time algorithms for circuits are either not easy to
% generalise or otherwise resist generalisation entirely. This follows from the
% fact that a circuit automorphism needs to preserve more structure when acting on
% non-symmetric gates than when acting on symmetric gates. As a result, while many
% symmetry-related circuit properties are polynomial-time decidable for circuits
% with symmetric gates, we have shown that these same properties are at least as
% hard as the graph-isomorphism for circuits with non-symmetric gates. This is a
% problem as the polynomial-time decidability of these properties is essential for
% our translation from circuits into formulas.

% As such, instead of working with general rank-circuits we have restricted our
% attention to \emph{transparent} circuits, and proved our characterisation in
% terms of $\PT$-uniform families of transparent symmetric rank-circuits. As we
% have shown in Section~\ref{sec:transparent}, all of the relevant properties are
% polynomial-time decidable for transparent circuits. Since all circuits with
% symmetric gates are transparent, we might think of transparent circuits as an
% alternative, more `algorithmically accessible', generalisation of the circuits
% studied by Anderson and Dawar. However, while this restriction to transparent
% circuits aids in our translation from circuits to formulas, it complicates our
% translation in the other direction. Indeed, it is not hard to see that the
% conventional approach for translating formulas of $\FO$ (or $\FPC$, or other
% extensions of $\FO$) into $\PT$-uniform families of circuits does not, in
% general, produce a family of transparent circuits. In order to address this we
% defined a novel translation from formulas of $\FPR$ into equivalent
% $\PT$-uniform families of transparent symmetric rank-circuits. This translation
% is defined in two steps. First, we defined a translation from $\FPR$ to
% $\PT$-uniform families of bounded-width formulas of $\FOrk$ and, second, we
% defined a translation from these families of formulas into $\PT$-uniform
% families of transparent symmetric rank-circuits. This translation, along with
% the main result of the paper, gives us an equivalence between the expressive
% power of three formalisms, which we summarise as follows:

@@

In short, we can represent the proof of our characterisation through
the three equivalences in this triangle.

\begin{center}
  \begin{tikzcd}
    \FPR \ar[r, equal] \ar[dr, equal] & \parbox{0.35\textwidth}{Uniform
      families of
      bounded-width $\FOrk$ formulas}  \ar[d, equal]\\
    & \parbox{0.38\textwidth}{Uniform families of transparent symmetric
      rank-circuits}
  \end{tikzcd}
\end{center}

This highlights another interesting aspect of our result.  The first
translation, of $\FPR$ to uniform families of $\FOrk$ formulas was
given in~\cite{Dawar09logicswith} and used there to establish arity
lower bounds.  However, this was for a weaker version of the rank
logic rather than the strictly more expressive one defined by
Gr\"{a}del and Pakusa~\cite{GradelP15a}.  The fact that we can
complete the cycle of equivalences with the more powerful logic
demonstrates that the definition of Gr\"{a}del and Pakusa is the
``right'' formulation of $\FPR$.

% We already have a very similar result from from Anderson and
% Dawar~\cite{AndersonD17} and Otto~\cite{Otto1996}:

% \begin{center}
%   \begin{tikzcd}
%     \FPC \ar[r, equal] \ar[dr, equal] & \parbox{0.35\textwidth}{$\PT$-uniform
%       families of
%       bounded-width $\FOc$-formulas}  \ar[d, equal]\\
%     & \parbox{0.35\textwidth}{$\PT$-uniform families of symmetric circuits with
%       symmetric gates}
%   \end{tikzcd}
% \end{center}

% These two sets of equivalences each give us three distinct formalisms equivalent
% to $\FPR$ and $\FPC$, respectively. Importantly, each of these three formalisms
% for $\FPC$ may be extended by a rank in quite distinct ways. We can extend
% $\FOc$ by rank quantifiers and define $\FOrk$, we can add rank gates to
% symmetric circuits with symmetric gates and define transparent symmetric
% rank-circuits, and we can extend $\FPC$ by a rank operator and define $\FPR$. We
% might expect these three quite different extensions to result in three
% formalisms with quite different expressive power. However, the main result of
% this paper along with the translation we discussed above, gives us that these
% three extensions are indeed all equivalent. This provides evidence in favour of
% the proposition that $\FPR$ is indeed the canonical, and most natural, extension
% of $\FPC$. This is in contrast with the other rank logic in the literature, the
% one introduced by Dawar et al.~\cite{Dawar09logicswith}, which was shown by
% Gr\"{a}del and Pakusa~\cite{GradelP15a} to be strictly less expressive than
% $\FPR$, and hence less expressive than $\PT$-uniform families of bounded-width
% $\FOrk$ and $\PT$-uniform families of transparent symmetric rank-circuits.

@@

\subsection*{Future Work}
There are many directions of work suggested by the methods and results
developed in this paper.  First of all, there is the question of
transparency.  We introduce it as a technical device that enables our
characterisation to go through.  Could it be dispensed with?  Or are $\PT$-uniform
families of transparent symmetric rank-circuits strictly weaker than
families without the restriction of transparency?  Of course a
positive answer to the latter question would require separating $\FPR$
from $\PT$.

The framework we have developed for  working with circuits with
structured inputs is very general and not specific to rank gates.
Indeed, the first use we make of linear algebra is in the proof of
Lemma~\ref{lem:rank-triple-equivilence}.  It would be interesting to
apply this framework to other logics.  It appears to be as general a way of
extending the power of circuits as Lindstr\"om quantifiers are in the
context of logic.  We would like to develop this link
further, perhaps for specific quantifiers such as $\FP$ extended by an operator
that expresses the solubility of systems of equations over rings as in~\cite{DGHKP}

% In this paper we have developed a framework for working with circuits over
% non-symmetric bases. In fact, the first time we explicitly use make any
% reference to linear algebra is in the proof of
% Lemma~\ref{lem:rank-triple-equivilence}, right towards the end of the paper. It
% would be natural to ask if our circuit characterisation of $\FPR$ can be
% generalised in order to develop similar characterisations of other logics. It
% might be good idea to start with some other extension of $\FP$ by some operator
% and try to develop a similar circuit characterisation in terms $\PT$-uniform
% transparent symmetric circuits over an appropriate basis. There are numerous
% logics that are similar in definition to $\FPR$ but which are not known to have
% the same expressive power as $\FPR$, for example $\FP$ extended by an operator
% that expresses the solubility of systems of equations over rings (rather than
% fields). A first step might be to develop a circuit characterisation of one of
% these logics.

@@

At the moment, we have little by way of methods for proving
inexpressibility results for $\FPR$, whether we look at it as a logic
or in the circuit model.  The logical formulation lays emphasis
on some parameters (the number of variables, the arity of the
operators, etc.) which we can treat as resources against which to
prove lower bounds.  On the other hand, the circuit model brings to
the fore other, more combinatorial, parameters.  One such is the
fan-in of gates and a promising and novel approach is to try and prove lower bounds for symmetric circuits with gates with bounded fan-in.
We might ask if it is possible to compute $\AND[3]$ using a symmetric circuit
with gates that have fan-in two.  Perhaps we could also combine the
circuit view with lower-bound methods from logic, such as pebble games.
 Dawar~\cite{Dawar2016} has shownn how the bijection games of
 Hella~\cite{Hella19961} can be used directly to prove lower
 bounds for symmetric without reference to the logic.  We also have pebble
 games for $\FPR$~\cite{DawarH2012}, and it would be interesting to
 know if we can use these on circuits and how the combinatorial
 parameters of the circuit interact with the game.

Finally, we note that some of the interesting directions on the
interplay between logic and symmetric circuits raised
in~\cite{AndersonD17} remain relevant.  Can we relax the symmetry
condition to something in between requiring invariance of the circuit
under the full symmetric group (the case of symmetric circuits) and
requiring no invariance condition at all?  Can such restricted
symmetries give rise to interesting logics in between $\FPR$ and
$\PT$?  It also remains a challenge to find a circuit characterisation
of  $\CPTC$.  Could the general framework for non-symmetric gates we
have developed here help in this respect?

% We have shown that the $\PT$-uniform families of transparent symmetric
% rank-circuits (resp. circuits with symmetric gates) have the same expressive
% power as $\FPR$ (resp. $\FPC$), but we recall that if we remove the symmetry
% restriction these circuit models express exactly those queries in $\PT$. A
% natural next step, also suggested by Anderson and Dawar, would be try and weaken
% the symmetry condition. One way this might be done would be to require that only
% permutations from a particular subgroup of the symmetric group extend to
% automorphisms of the circuit. It is easy to see that if we take this subgroup to
% be the trivial subgroup then this nullifies the the symmetry requirement, and
% the circuit model has the same expressive power as $\PT$, while if we take this
% subgroup to be the entire symmetric group then we get back the full symmetry
% requirement and hence our model has the same expressive power as the logic. As
% such, this approach can be used to interpolate between the logic and $\PT$.

% There is a rather easy way to develop a circuit characterisation for any logic
% or complexity class with complete problems under first-order reductions. All
% that is required is to extend the basis with the family of functions that decide
% a complete problem and consider $\PT$-uniform families of transparent symmetric
% circuits over this basis. We did not take this approach in our characterisation
% of $\FPR$, preferring to use a more `natural' basis that more closely resembles
% the actual definition of the logic. We should like to ask if there is a similar
% circuit characterisation of $\CPTC$. One might start with bounded-rank fragments
% of $\CPTC$ and see if these programs can be translated into circuits. We should
% note that it does seem as though the symmetry inherent in $\CPTC$ programs is in
% some sense different from the symmetry of symmetric circuits. It might be more
% natural then to ask if $\CPTC$ can be characterised using families of circuits
% with a weakened symmetry requirement.
